### 2021-12-07 input/proposal.tex
proof FB + algo

\begin{definition}[\label{def:forwarding}Forwarding]
  A \process $x$ \underline{f}orwards a message $m$ by
  \underline{s}ending it to its neighbors with metadata that depends
  on its local knowledge: $f_x(m) \implies \forall \langle x, y\rangle
  \in E: s_{xy}(m \oplus \sigma_x)$.
\end{definition}

@@

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to eventually ($\eventually$) reach
consistent partitioning, each source $s$ must send notifications
$\alpha_s$ to all \processes that are closer to it than any other
source, along with enough control information to allow them to decide
of their closest source.

\begin{theorem}[\label{theo:fb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually \textup{CP}$]
    % 
    Assuming reliable communication links ($s_{xy}(m) \iff \eventually
    r_{yx}(m)$), and a total order on messages ($<$), \processes
    eventually reach consistent partitioning if each \process delivers
    its best message among \underline{r}eceived messages $R_x$, and
    forwards its best message among \underline{d}elivered messages
    $D_x$.
    % , \ie
    % the message $\alpha_s^d$ with the smallest distance $d$ to the
    % source $s$
    % ($(\alpha_s^d = \min R_x) \implies (d_x(\alpha_s^d) \wedge
    % f_x(\TODO{\alpha_s^d}))$).
\end{theorem}

\begin{proof}
  % When there are no sources, no \process receives, nor delivers hence
  % forwards any message. \Processes do not belong to any partition.

  \noindent \TODO{use total order ($<$) and sets $R_p D_p$} Whenever a
  \process $s$ becomes a source, it delivers its own message
  $d_s(\alpha_{s})$. Whatever its set of received messages, its own
  partition $P_s$ becomes and remains its closest partition
  $\alpha_{s}^{|\Pi_{ss}|} = \min D_s$ since $\forall x \neq s:
  |\Pi_{s s}| < |\Pi_{s x}|$.
  %
  \noindent Such a source $s$ forwards $\alpha_{s}$ to its neighbors.
  Since communication links are reliable, neighboring \processes
  eventually receive such a notification
  $\forall \langle s, y \rangle \in E \iff \eventually
  r_y(\alpha_{s}^{d})$. Most importantly, whatever the order of
  received messages, every \process $q$ in this neighborhood such that
  there exists no closer source $s'$ than the received one $s$,
  delivers it, since $w_{sq} < |\Pi_{s'q}|$.
  %
  \noindent They forward it and their respective neighbors eventually
  receive it. By transitivity, the message originating from $s$
  reaches all \processes belonging to $P_s$ at least through their
  shortest paths: $\smash{\forall q' \in V, s, s' \in S: |\Pi_{s q'}| <
  |\Pi_{s' q'}| \implies \eventually d_{q'}(\alpha_{s}^{|\Pi_{sq'}|})}$.
  Since there exists only one best sum of weights per \process that
  can never be retracted, \processes eventually reach consistent
  partitioning.
  %% /!\ not equivalence for there exists other implementations.
  %% \item [$CP \implies BEF$:] By contradiction, if a \process $q \in
  %%   \Pi_{sp} = [s, \ldots, q, q', \ldots, p]$ with $s, q, p \in P_s$
  %%   does not forward its received $\alpha_s^{|\Pi_{sq}|}$, then
  %%   following \processes from $q'$ to $p$ may mistake another
  %%   partition for their  because it needs the weight $W_{pq}$.
  %% \end{asparadesc}
\end{proof}

@@

Algorithm~\ref{algo:add} shows the instructions that implement
forwarding of best to reach consistent partitioning when
\begin{inparaenum}[(i)]
\item weights are scalar values piggybacked in messages and
  accumulated over forwarding,
\item \processes only add new partitions to the system,
\item and \processes never join, crash nor leave the system.
\end{inparaenum}
For the sake of simplicity and to remove obvious loops, a \process
does not forward a notification to its sender
(Line~\ref{line:accumulator}).  Figure~\ref{fig:add} illustrates the
behavior of this algorithm on a system comprising 4 \processes $a$,
$b$, $c$, and $d$. \Process~$a$ and \Process~$d$ become the sources of
their partition.  They \NAMEB a notification \TODO{\underline{a}dd
  message}: $\alpha_a^0$ and $\alpha_d^0$. They initialize their own
state with the lowest possible bound $0$ (see
Line~\ref{line:lowestbound}), and send a message to each of their
neighbors by accumulating the corresponding edge weight (see
Line~\ref{line:accumulator}). In Figure~\ref{fig:addC}, \Process~$b$
receives $\alpha_{d}^{1}$. Since it improves its own partition
distance, it keeps it and forwards it to its neighbors. In
Figure~\ref{fig:addD}, $b$ discards $\alpha_{a}^{2}$, for it does not
improve its partition distance. \Processes $c$ and $d$ will never
acknowledge that Source~$a$ exists. Ultimately, \processes discard
last transiting messages. Despite the obvious lack of traffic
optimization, the system reaches consistent partitioning.

>>>>

\begin{definition}[\label{def:forwarding}Forwarding]
  A \process $x$ \underline{f}orwards a message $m$ by
  \underline{s}ending it to its neighbors $y$ with metadata that
  depends on its local knowledge: $f_x(m) \implies \forall \langle x,
  y\rangle \in E: s_{xy}(m \oplus \sigma_x)$.
\end{definition}

@@

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to eventually ($\eventually$) reach
consistent partitioning, each source $s$ must send
\texttt{\underline{A}dd} notifications $\alpha_s$ to all \processes
that are closer to it than any other source, along with enough
metadata to allow them to decide of their closest source. A simple
implementation consists in ensuring that each \process $x$ eventually
receives $\alpha_s^{|\Pi_{sx}|}$ where $\Pi_{sx}$ is the \emph{best}
path from any source to it; the value $|\Pi_{sx}|$ being built over
forwarding along this path.

\begin{theorem}[\label{theo:fb}\underline{F}orwarding of \underline{B}est (FB)
    $\implies\eventually \textup{CP}$]
    % 
    Assuming that each \process $x$ stores its outgoing weights
    ($\forall \langle x, y \rangle \in E: \pi_{xy} \in \sigma_x$), a
    total order on messages based on weights ($m^d_s < m^{d'}_{s'}$
    when $d < d' \vee (d = d' \wedge s< s')$), reliable communication
    links ($s_{xy}(m) \iff \eventually r_{yx}(m)$), \processes eventually reach
    consistent partitioning if each \process delivers its best (\ie
    smallest) message among \underline{r}eceived messages $R_x$, and
    forwards its best message among \underline{d}elivered messages
    $D_x$ such that $f_x(\alpha^d_s) \implies \forall \langle x, y
    \rangle \in E: s_{xy}(\alpha_s^d \oplus ^{w_{xy}})$, \ie by
    accumulating the respective weight of used edges.
\end{theorem}

\begin{proof}
  Whenever a \process~$s$ becomes a source, it broadcasts hence
  delivers its own message $d_s(\alpha_{s}^{|\pi_{ss}|})$. Whatever
  its set of received messages, it acknowledges that it belongs to its
  own partition $P_s$ since $\alpha_{s}^{|\pi_{ss}|} = \min D_s$ and
  it remains forever since $|\pi_{ss}|$ is its greatest lower bound.
  \noindent Such a source forwards its notification to its neighbors.
  Every neighbor eventually receives its notification since
  communication links are reliable. Whatever the order of received
  messages $R_x$ at neighboring \process~$x$, total order ensures that
  it delivers notifications $\alpha_s^{d}$ when $d= |\pi_{ss}| +
  w_{sx} = \min |\pi_{s'x}|$ where $\min |\pi_{s'x}|$ is the lightest
  weight of $x$ to any source $s'$.
  \noindent Among these neighbors, at least those that fulfill the
  latest condition forward their respective notification.  By
  transitivity, the message originating from $s$ reaches all
  \processes that belong to $P_s$ at least through their respective
  lightest path: $\small\smash{\forall y \in V, s, s' \in S: \min
    |\pi_{s y}| < \min |\pi_{s' y}| \implies \eventually
    d_{y}(\alpha_{s}^{\min |\pi_{sy}|})}$. When the system becomes
  quiescent, \ie no \process becomes source anymore, every \process
  eventually acknowledges the partition it belongs to, \ie \processes
  eventually reach consistent partitioning together. In addition, the
  protocol terminates: a \process never delivers hence forwards a
  message after it received, delivered, and forwarded the message of
  its closest source from its lightest path.
\end{proof}

@@

Algorithm~\ref{algo:add} shows the few instructions that implement a
\emph{reactive} forwarding of best to reach consistent partitioning in
a static network where \processes never join, crash, nor leave the
system. As soon as a \process receives a better message, it delivers
and forwards it. The trade-off consists in decreasing termination time
at cost of increased traffic. For the sake of simplicity and to remove
obvious loops, a \process does not forward a notification to its
sender (Line~\ref{line:accumulator}). Figure~\ref{fig:add} illustrates
the behavior of this algorithm on a system comprising 4 \processes
$a$, $b$, $c$, and $d$. Both $a$ and $d$ become sources.  They \NAMEB
notifications $\alpha_a^0$ and $\alpha_d^0$. They initialize their own
state with the lowest value $0$ (see Line~\ref{line:lowestbound}), and
send a message to each of their neighbors by accumulating the
corresponding edge weight (see Line~\ref{line:accumulator}). In
Figure~\ref{fig:addC}, $b$ receives $\alpha_{d}^{1}$. Since it
improves its own partition distance, it keeps it and forwards it to
its neighbors. In Figure~\ref{fig:addD}, $b$ discards
$\alpha_{a}^{2}$, for it does not improve its partition
distance. Neither $c$ nor $d$ will ever acknowledge that Source~$a$
exists. Ultimately, \processes discard last transiting
messages.
% Despite the obvious lack of traffic optimization,
The system reaches consistent partitioning.
